---
title: Mixed Model
authors: "Carina Huang, Evan Mularo, and Tallulah Bowden"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Setup

```{r}
# Load necessary packages
library(here)
library(tidyverse)
library(lubridate)
library(nlme)
library(gtsummary)
library(DHARMa)
library(lme4)

# Load necessary data
bleaching <- read.csv("Code/bahamas_florida_bleaching_subset.csv")
bleaching$Percent_Bleaching <- as.numeric(parse_number(bleaching$Percent_Bleaching))
bleaching$Temperature_Mean <- as.numeric(parse_number(bleaching$Temperature_Mean))
bleaching$Turbidity <- as.numeric(parse_number(bleaching$Turbidity))
bleaching$Distance_to_Shore <- as.numeric(parse_number(bleaching$Distance_to_Shore))

# Get rid of rows with NAs
bleaching <- bleaching %>%
  drop_na()

# Make percent coral bleaching appropriate for binomial regression
bleaching <- bleaching %>%
  mutate(Bleaching_pct = Percent_Bleaching/100)
```


### Step 1: Define your research question.

### Do sea surface temperature, distance to shore, turbidity, cyclone events, percent coral bleaching, and site ID significantly predict percent cover of coral bleaching?

### Step 3: Fit linear regression model without random effect and evaluate model diagnostics

```{r}
# Fit linear regression model
bleaching_model1 <- lm(Bleaching_pct ~ Temperature_Mean + Turbidity + Cyclone_Frequency + Distance_to_Shore,
                        data = bleaching)
# Examine the results and residuals
summary(bleaching_model1)
plot(bleaching_model1)

```

### Step 4: Fit linear regression model with random effect and evaluate model diagnostics

```{r}
# Re-fit standard linear regression
bleaching_model2 <- gls(Bleaching_pct ~ Temperature_Mean + Turbidity + Cyclone_Frequency + Distance_to_Shore, data = bleaching)

# Fit a mixed effects model with random (varying) intercept by plot
bleaching_mixedm1 <- glmer(Bleaching_pct ~ Temperature_Mean + Turbidity + Cyclone_Frequency + Distance_to_Shore + (1|Site_ID), data = bleaching,
                  family = binomial)

# Compare the two model structures
AIC(bleaching_model2, bleaching_mixedm1)

# Examine model residual plot
plot(bleaching_mixedm1)

# Need to examine QQ plot separately
qqnorm(bleaching_mixedm1)

# Examine model results
summary(bleaching_mixedm1)
```

# Visualize

```{r}
# Sample report visualization
fig1 <- ggplot(bleaching, aes(x = Turbidity, y = Bleaching_pct)) +
 # geom_boxplot(width = 0.2) +
  geom_point(aes(color = Site_ID), size = 3,  alpha = 0.5) +
  #scale_color_manual(values = c("#ECBD95", "#FFAA00", "#D46F10", "#4CA49E", 
   #                             "#69B9FA", "#4B8FF7", "grey30")) +
  labs(x = "Turbidity",
       y = "PCT Bleaching") +
  theme_bw()
fig1
```

